Dependent Object Types

### 2012-09-03 doc/candidacy/report.tex
syntax-directed does not imply decidable
tribe rev

Path-dependent types are a key concept in the DOT calculus, and a
unifying concept of the three papers we choose to study. A {\it
  dependent type} is a type which depends on a term. A {\it
  path-dependent type} is a limited form of dependent type, in which a
type depends on a path: a chain of immutable fields or
stable values. Path-dependent types arise when terms, usually objects, can
have types as members. Abstract type members, together with explicit
selftypes and modular mixin composition have been identified as key
abstractions for the construction of reusable components~\cite{SCA}.

@@

In~\cite{tribe}, Clarke et al. present a calculus for virtual
classes. Virtual classes allow {\it families} of classes to be
inherited, rather than just single classes. Classes are lexically
nested inside other classes. When a class is inherited, its nested
inner classes are inherited along with its methods and fields. {\it
  Further binding} enables the nested classes to be further
extended. Path-dependent types arise naturally to distinguish objects
coming from different families. In Tribe, path types can depend
simultaneously on both classes and objects. Additionally, paths can
use an \code{out} field to move from an object to the object which
surrounds it. This \code{out} feature enables ubiquitous access to an
object's family without the need to drag around family arguments.

@@

In our own research, we have been exploring a new core calculus for
path-dependent types, the Dependent Object Types (DOT) calculus, which
could serve as a new type-theoretic foundation for Scala and languages
like it. The three calculi presented demonstrate that path-dependent
types arise in diverse settings, and thus, it is valuable to identify
a core calculus which captures their essence. In DOT, we are not
concerned with modeling orthogonal concepts such as inheritance and
mixin composition. In contrast to the three previous calculi,
type-checking in DOT is decidable, since all the judgments are
syntax-directed (in the current design).

>>>>

Path-dependent types are a key concept in the DOT calculus, and a
unifying concept of the three papers we choose to study. A {\it
  dependent type} is a type which depends on a term. A {\it
  path-dependent type} is a limited form of dependent type, in which a
type depends on a path: a chain of immutable fields or
stable values. Path-dependent types arise when terms, usually objects, can
have types as members. Abstract type members, together with explicit
selftypes and modular mixin composition, have been identified as key
abstractions for the construction of reusable components~\cite{SCA}.

@@

In~\cite{tribe}, Clarke et al. present a calculus for virtual
classes. Virtual classes allow {\it families} of classes to be
inherited, rather than just single classes. Classes are lexically
nested inside other classes. When a class is inherited, its nested
inner classes are inherited along with its methods and fields. {\it
  Further binding} enables the nested classes to be further
extended. Path-dependent types arise naturally to distinguish objects
coming from different families. Yet, thanks to {\it family
  polymorphism}, code written for one family also works for extension
of that family. In Tribe, path types can depend simultaneously on both
classes and objects. Additionally, paths can use an \code{out} field
to move from an object to the object which surrounds it. This
\code{out} feature enables ubiquitous access to an object's family
without the need to drag around family arguments.

@@

In our own research, we have been exploring a new core calculus for
path-dependent types, the Dependent Object Types (DOT) calculus, which
could serve as a new type-theoretic foundation for Scala and languages
like it. The three calculi presented demonstrate that path-dependent
types arise in diverse settings, and thus, it is valuable to identify
a core calculus which captures their essence. In DOT, we are not
concerned with modeling orthogonal concepts such as inheritance and
mixin composition.
